### 2017-01-26 paper.Rmd
Integration of cost data and first draft paper

---
title: "Cost-effectiveness of a malaria elimination campaign"
author: "Laia Cirera, Joe Brew, Elisa Sicuri"
date: "26 de enero de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
library(readxl)
library(tidyverse)

# Read in Laias cost data
costs <- read_excel("public_data/costs.xlsx")

# Get yearly malaria cases for magude

# Create a column for % severe malaria
# This will be the number of cases of ...

# Read in incidence data
bes <- read_csv("data/cleaned/cases.csv")
```

# Introduction

The average incidence of malaria for our years of observation was `r round(mean(bes$pk, na.rm=TRUE),digits=2)`

# Methodology

WE analyzed data from bla bla bla

# Results

Our results show that the effectiveness of MALTEM's interventions is as follows:

```{r}
plot(x=costs$year , y=costs$`Case fatality rate`, type="l",
     xlab="YEAR", ylab="rate", main="Case fatality rate ACROSS TIME",
     col="blue")
```

And here is another figure.

```{r, fig.height = 8}

ggplot(data = bes,
       aes(x = date,
           y = pk)) +
  geom_line(aes(color = age)) +
  facet_wrap(~district) +
  labs(x = 'Date',
       y = 'Incidence (cases per 1,000)',
       title = 'Malaria incidence over time by district',
       subtitle = 'Cases per 1,000 inhabitants') +
  scale_color_manual(name = 'Age',
                     values = c('darkgreen', 'darkorange'))
```

# Discussion