### 2017-07-16 reports/misc_analysis/misc_analysis.Rmd
questions for london

---
title: "MALTEM effectiveness analysis"
author: "Brew, Cirera, Sicuri, Thomas"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

# Packages 
library(databrew)
library(cism)
library(dplyr)
library(ggplot2)
library(ggmap)
library(gsheet)
library(readr)

# Options for this document (ie, whether to show code, errors, etc.)
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE,
               eval = TRUE,
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
# Read in wide weather data
# source('../../weather_functions.R', chdir = TRUE)
source('../../master.R', chdir = TRUE)

# Define a theme
theme_maltem <- ggthemes::theme_hc
```

# Visual overview

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_line(alpha = 0.8,
            color = 'darkred') +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  facet_wrap(~district) +
  theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, age_group, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = age_group,
           color = age_group)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('darkorange', 'darkgreen')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~district) +
  # theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem()

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place, age_group) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~age_group)
```

# When did the decline begin in Magude?

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  group_by(date) %>%
  summarise(magude_to_other = p[place == 'Magude'] / 
              p[place == 'Other districts'] * 100)

g <- ggplot(data = x,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  geom_hline(yintercept = 100,
             col = 'darkred',
             alpha = 0.6,
             lty = 2) +
  labs(x = 'Date',
       y = 'Percent') +
  labs(title = 'Monthly incidence: Magude as % of other districts',
       subtitle = 'All age groups') +
  theme_maltem()

g + xlim(as.Date('2013-01-01'), max(x$date))

```

# Magude vs. Manhiça

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(district %in% c('MAGUDE', 'MANHICA')) %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  
                        ifelse(district == 'MANHICA', 'Manhiça', NA))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. Manhiça',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

```

# Weather

## Precipitation

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

g <- ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = precipitation,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = precipitation,
           group = district)) +
  labs(x = 'Date',
       y = 'Milimeters',
       title = 'Precipitation: Magude vs. other districts',
       subtitle = 'Monthly average precipitation, Magude in red') +
  theme_maltem()
g 
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')

y <- x %>%
  group_by(date) %>%
  summarise(magude_to_other = mean(precipitation[district == 'MAGUDE']) / 
              mean(precipitation[district != 'MAGUDE']) * 100)
g <- 
  ggplot(data = y,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  theme_maltem() +
  labs(x = 'Date',
       y = 'Percent',
       title = 'Average monthly precipitation: Magude vs other',
       subtitle = 'Magude as percentage of monthly mean') +
  geom_hline(yintercept = 100,
             color = 'darkred',
             lty = 2,
             alpha = 0.7) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3)
g
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')
```

## Temperature

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = temp,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = temp,
           group = district)) +
  labs(x = 'Date',
       y = 'Degrees (celcius)',
       title = 'Average temperature: Magude vs. other districts',
       subtitle = 'Monthly average, Magude in red') +
  theme_maltem() +
  xlim(as.Date('2014-01-01'),
               max(x$date))
```

## Is seasonality constant by year?

```{r}
x <- 
  df %>%
  group_by(year, month) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(year = factor(year))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = month,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  geom_point() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  labs(x = 'Month',
       y = 'Average incidence',
       title = 'Incidence by month, all districts',
       subtitle = 'Wide variability in monthly incidence') +
  theme_maltem()
```

## Does 4 week long-and-lagged precipitation predict incidence?

Yes, but not in the way we'd expect. When we adjust for the fixed effect of month, we get a negative correlation between lagged precipitation and incidence.

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(date <= '2015-06-01') %>%
  mutate(year_month = as.Date(paste0(year, '-', month, '-01'))) %>%
  arrange(date) %>%
  group_by(district) %>%
  mutate(lag4 = lag(precipitation, n = 4, default = NA),
            lag5 = lag(precipitation, n = 5, default = NA),
            lag6 = lag(precipitation, n = 6, default = NA),
            lag7 = lag(precipitation, n = 7, default = NA)) %>%
  mutate(p = sum(cases) / sum(population) * 1000) %>%
  mutate(n_nas = ifelse(is.na(lag4), 1, 0) + 
           ifelse(is.na(lag5), 1, 0) + 
           ifelse(is.na(lag6), 1, 0) + 
           ifelse(is.na(lag7), 1, 0)) %>%
  mutate(n_ok = 4 - n_nas) %>%
  mutate(precipitation_lagged = ifelse(is.na(lag4), 0, lag4) +
           ifelse(is.na(lag5), 0, lag5) +
           ifelse(is.na(lag6), 0, lag6) +
           ifelse(is.na(lag7), 0, lag7) / 
           n_ok) %>%
  filter(!is.na(precipitation_lagged),
         !is.na(p))

fit <- lm(p ~ precipitation_lagged + 
            factor(month), data = x)
broom::tidy(fit)
```

# Proximity to South Africa

# Questions

1. What are the mathematics of the method by which we assign weights to synthetic control districts? Are we concerned that so few districts get any weight?