### 2017-07-19 reports/misc_analysis/misc_analysis.Rmd

---
title: "MALTEM effectiveness analysis"
author: "Brew, Cirera, Sicuri, Thomas"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

# Packages 
library(databrew)
library(cism)
library(dplyr)
library(ggplot2)
library(ggmap)
library(gsheet)
library(readr)

# Options for this document (ie, whether to show code, errors, etc.)
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE,
               eval = TRUE,
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
# Read in wide weather data
# source('../../weather_functions.R', chdir = TRUE)
source('../../master.R', chdir = TRUE)

# Define a theme
theme_maltem <- ggthemes::theme_hc
```

# Visual overview

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_line(alpha = 0.8,
            color = 'darkred') +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  facet_wrap(~district) +
  theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, age_group, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = age_group,
           color = age_group)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('darkorange', 'darkgreen')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~district) +
  # theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem()

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place, age_group) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~age_group)
```

# When did the decline begin in Magude?

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  group_by(date) %>%
  summarise(magude_to_other = p[place == 'Magude'] / 
              p[place == 'Other districts'] * 100)

g <- ggplot(data = x,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  geom_hline(yintercept = 100,
             col = 'darkred',
             alpha = 0.6,
             lty = 2) +
  labs(x = 'Date',
       y = 'Percent') +
  labs(title = 'Monthly incidence: Magude as % of other districts',
       subtitle = 'All age groups') +
  theme_maltem()

g + xlim(as.Date('2013-01-01'), max(x$date))

```

# Magude vs. Manhiça

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(district %in% c('MAGUDE', 'MANHICA')) %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  
                        ifelse(district == 'MANHICA', 'Manhiça', NA))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. Manhiça',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

```

# Weather

## Precipitation

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

g <- ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = precipitation,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = precipitation,
           group = district)) +
  labs(x = 'Date',
       y = 'Milimeters',
       title = 'Precipitation: Magude vs. other districts',
       subtitle = 'Monthly average precipitation, Magude in red') +
  theme_maltem()
g 
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')

y <- x %>%
  group_by(date) %>%
  summarise(magude_to_other = mean(precipitation[district == 'MAGUDE']) / 
              mean(precipitation[district != 'MAGUDE']) * 100)
g <- 
  ggplot(data = y,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  theme_maltem() +
  labs(x = 'Date',
       y = 'Percent',
       title = 'Average monthly precipitation: Magude vs other',
       subtitle = 'Magude as percentage of monthly mean') +
  geom_hline(yintercept = 100,
             color = 'darkred',
             lty = 2,
             alpha = 0.7) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3)
g
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')
```

## Temperature

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = temp,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = temp,
           group = district)) +
  labs(x = 'Date',
       y = 'Degrees (celcius)',
       title = 'Average temperature: Magude vs. other districts',
       subtitle = 'Monthly average, Magude in red') +
  theme_maltem() +
  xlim(as.Date('2014-01-01'),
               max(x$date))
```

## Is seasonality constant by year?

```{r}
x <- 
  df %>%
  group_by(year, month) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(year = factor(year))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = month,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  geom_point() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  labs(x = 'Month',
       y = 'Average incidence',
       title = 'Incidence by month, all districts',
       subtitle = 'Wide variability in monthly incidence') +
  theme_maltem()
```

## Does 4 week long-and-lagged precipitation predict incidence?

Yes, but not in the way we'd expect. When we adjust for the fixed effect of month, we get a negative correlation between lagged precipitation and incidence.

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(date <= '2015-06-01') %>%
  mutate(year_month = as.Date(paste0(year, '-', month, '-01'))) %>%
  arrange(date) %>%
  group_by(district) %>%
  mutate(lag4 = lag(precipitation, n = 4, default = NA),
            lag5 = lag(precipitation, n = 5, default = NA),
            lag6 = lag(precipitation, n = 6, default = NA),
            lag7 = lag(precipitation, n = 7, default = NA)) %>%
  mutate(p = sum(cases) / sum(population) * 1000) %>%
  mutate(n_nas = ifelse(is.na(lag4), 1, 0) + 
           ifelse(is.na(lag5), 1, 0) + 
           ifelse(is.na(lag6), 1, 0) + 
           ifelse(is.na(lag7), 1, 0)) %>%
  mutate(n_ok = 4 - n_nas) %>%
  mutate(precipitation_lagged = ifelse(is.na(lag4), 0, lag4) +
           ifelse(is.na(lag5), 0, lag5) +
           ifelse(is.na(lag6), 0, lag6) +
           ifelse(is.na(lag7), 0, lag7) / 
           n_ok) %>%
  filter(!is.na(precipitation_lagged),
         !is.na(p))

fit <- lm(p ~ precipitation_lagged + 
            factor(month), data = x)
DT::datatable(broom::tidy(fit))
```

# Proximity to South Africa

We calculate the distance from the centroid of each district to the South African border.

```{r}
# Interpolate for our locations
library(sp)
x <- cism::moz2
x <- x[x@data$NAME_1 %in% c('Maputo', 'Gaza'),]
coords <- data.frame(coordinates(x))
names(coords) <- c('x', 'y')
coords$district <- x@data$NAME_2

# Standardize names
coords$district <- toupper(coords$district)
coords$district[coords$district == 'CHÓKWÈ'] <- 'CHOKWE'
coords$district[coords$district == 'GUIJÁ'] <- 'GUIJA'
coords$district[coords$district == 'XAI-XAI'] <- 'XAI-XAI DISTRICT'
coords$district[coords$district == 'MANHIÇA'] <- 'MANHICA'
coords$district[coords$district == 'MATUTUÍNE'] <- 'MATUTUINE'
coords$district[coords$district == 'MANDLAKAZI'] <- 'MANJACAZE'

# Add a row for matola
matola <- data_frame(x = 32.46580,
                     y = -25.92391,
                     district = 'MATOLA')
coords <- bind_rows(coords, 
                    matola)
# Add a row for xai-xai city
xxc <- coords %>% filter(district == 'XAI-XAI DISTRICT') %>%
  mutate(district = 'XAI-XAI CITY')
coords <- bind_rows(coords,
                    xxc)
# Make spatial
coords <- coords %>%
  mutate(lng = x,
         lat = y)
coordinates(coords) <- ~x+y
proj4string(coords) <- proj4string(moz3)

# Get South Africa
sa <- cism::africa
sa <- sa[sa@data$COUNTRY == 'South Africa',]

# Get projected versions
sa_proj <- spTransform(sa,
                    CRS("+init=epsg:3347"))
coords_proj <- spTransform(coords,
                    CRS("+init=epsg:3347"))

# Get distance to south Africa
coords@data$distance_to_south_africa <- NA
for (i in 1:nrow(coords)){
  this_distance <- rgeos::gDistance(coords_proj[i,], sa_proj)
  coords@data$distance_to_south_africa[i] <- this_distance / 10000
}

afr <- cism::africa
afr <- afr[afr@data$COUNTRY %in% c('South Africa', 'Mozambique'),]
plot(afr, col = adjustcolor(c('blue', 'green'), alpha.f = 0.3))
points(coords, col = 'red',
       pch = '.',
       cex = 2)
```

The distances can be seen in the below table.

```{r}
DT::datatable(coords@data %>% dplyr::select(district, distance_to_south_africa) %>%
                mutate(distance_to_south_africa = round(distance_to_south_africa, digits = 2)) %>%
  rename(District = district,
         KM = distance_to_south_africa)) 
```

Visually, we assess the hypothesis that the decline over time is steeper for those districts bordering South Africa through a series of charts.

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  left_join(coords@data) %>%
  group_by(district, year, distance_to_south_africa) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = distance_to_south_africa,
           y = p)) +
  geom_point(alpha = 0.7) +
  geom_smooth(n = 5, span = 10, se = FALSE,
              color = 'darkred',
              alpha = 0.5,
              lty = 1) +
  facet_wrap(~year) +
  theme_maltem() +
  labs(x = 'Kilometers to South Africa',
       y = 'Weekly incidence per K',
       title = 'Protective effect of South African proximity',
       subtitle = 'One point per district per year')

```

As an alternative to the above, we can visualize the same data with the same aesthetics, but overlaying each year.

```{r}
library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = distance_to_south_africa,
           y = p,
           color = factor(year))) +
  geom_point(alpha = 0.7) +
  geom_smooth(n = 5, span = 10, se = FALSE,
              alpha = 0.5,
              lty = 1) +
  theme_maltem() +
  labs(x = 'Kilometers to South Africa',
       y = 'Weekly incidence per K',
       title = 'Protective effect of South African proximity',
       subtitle = 'One point per district per year') +
  scale_color_manual(name = 'Year',
                     values = cols)
```

Could the proximity to South Africa partly explain the pre-MALTEM decline in incidence in Magude? We'll examine this by modeling incidence as a linear function of district, distance, the interaction of year and distance, and month, in effect constructing a "counter-factual" for Magude that takes into account the secular decline in incidence associated with proximity to South Africa. For the model's purpose, we take into account only those 4 years from 2012-06-01 through 2016-06-01 (so as to not include the effect of MALTEM). Our model formula predicts incidence as a function of the interactions of season/year and distance/year, with the fixed effect of month.

```{r}
model_data <- 
  df %>%
  filter(disease == 'MALARIA') %>%
  left_join(coords@data) %>%
  mutate(season = ifelse(month %in% 1:4, 'rainy', 'not-rainy')) %>%
  mutate(month = factor(month)) %>%
  group_by(date, year, 
           season,
           month,
           district, distance_to_south_africa) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup

fit <- lm(p ~ year*log(distance_to_south_africa)  + year + season*year + month, 
          data = model_data %>%
            filter(date <= '2016-06-01',
                   date >= '2012-06-01'))

DT::datatable(broom::tidy(fit))

model_data$predicted <- predict(fit, model_data)

ggplot() +
  geom_line(data = model_data %>%
         filter(district == 'MAGUDE'),
       aes(x = date,
           y = predicted),
       alpha = 0.8) +
  geom_line(data = model_data %>%
         filter(district == 'MAGUDE'),
       aes(x = date,
           y = p),
       color = 'red',
       alpha = 0.8) +
  labs(title = 'Weekly malaria incidence in Magude',
       subtitle = 'Observed (red) vs. predicted (black)',
       x = 'Date',
       y = 'Cases per 1,000') +
  theme_maltem()
```

# Effect of insecticide-treated bednets (ITN)

At first glance, ITN coverage appears associated with lower malaria incidence

```{r}
ggplot(data = df %>% filter(disease == 'MALARIA'),
       aes(x = itn_coverage,
           y = pk)) +
  geom_point(alpha = 0.3) +
  theme_maltem() +
  labs(x = 'ITN coverage',
       y = 'Weekly incidence per 1,000',
       title = 'ITN and malaria',
       subtitle = 'One point per district, week, age-group') +
  scale_x_log10()
```

# Effect of indoor residual spraying (IRS)

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district, irs_coverage_people, weeks_since_last_irs_campaign_end) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(irs_time_since = ifelse(weeks_since_last_irs_campaign_end <= 10,
                                 '00-10',
                                 ifelse(weeks_since_last_irs_campaign_end <= 50,
                                        '11-50',
                                        ifelse(weeks_since_last_irs_campaign_end <= 99,
                                               '51-99',
                                               ifelse(weeks_since_last_irs_campaign_end > 100,
                                                      '99+',
                                                      NA))))) %>%
  filter(!is.na(weeks_since_last_irs_campaign_end)) %>%
  mutate(irs_coverage = ifelse(irs_coverage_people < 50,
                               '00-50%',
                               ifelse(irs_coverage_people <= 70,
                                      '51-70%',
                                      ifelse(irs_coverage_people <= 90,
                                             '71-90%',
                                             ifelse(irs_coverage_people > 90,
                                                    '91+%',
                                                    '00%')))))

ggplot(data = x,
       aes(x = weeks_since_last_irs_campaign_end,
           y = p)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~irs_coverage) +
  labs(x = 'Weeks since last IRS campaign end',
       y = 'Incidence per 1,000',
       title = 'Effect of IRS',
       subtitle = 'Faceted by coverage percentage') +
  theme_maltem()

```

# Incorporating more variables

Here's what happens when we model cases using ITN, IRS, time since IRS campaign, seasonality, etc.

```{r}
model_data <- 
  df %>%
  filter(disease == 'MALARIA') %>%
  left_join(coords@data) %>%
  mutate(season = ifelse(month %in% 1:4, 'rainy', 'not-rainy')) %>%
  mutate(month = factor(month)) %>%
  group_by(date, year, 
           season,
           month,
           irs_coverage_people,
           weeks_since_last_irs_campaign_end,
           itn_coverage,
           district, distance_to_south_africa) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup

fit <- lm(p ~ year*log(distance_to_south_africa)  + year + season*year  + month +  irs_coverage_people*weeks_since_last_irs_campaign_end + itn_coverage + district, 
          data = model_data %>%
            filter(date <= '2016-06-01',
                   date >= '2012-06-01'))

x <- model_data %>% filter(district == 'MAGUDE')
x$predicted <- predict(fit, x)

ggplot() +
  geom_line(data = x,
       aes(x = date,
           y = predicted),
       alpha = 0.8) +
  geom_line(data = x,
       aes(x = date,
           y = p),
       color = 'red',
       alpha = 0.8) +
  labs(title = 'Weekly malaria incidence in Magude',
       subtitle = 'Observed (red) vs. predicted (black)',
       x = 'Date',
       y = 'Cases per 1,000') +
  theme_maltem()

```

# Questions

1. What are the mathematics of the method by which we assign weights to synthetic control districts? Are we concerned that so few districts get any weight?

2. Given that proximity to South Africa seems to have an interactive effect with the decline in malaria incidence, should we include this in our model? If so, how?

3. Are we currently excluding Manhiça from our synthetic controls? If so, why?

4. What are we doing with IRS / ITN?

5. How are we treating secular trends (ie, near linear decline in springtime malaria in Magude for the 4 years from 2012-2015)?