### 2017-08-18 visualizations.Rmd
running figures locally

---
title: "Publication-quality visualizations for MALTEM effectiveness paper"
subtitle: "Reproducible code and output"
author: "Brew, Cirera, Sicuri, Thomas"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

What follows is the code used for the generation of publication quality figures. Note that certain dependencies' source code are not included in this document (for example, references to other `R` scripts); in these cases, all code is available at https://github.com/joebrew/maltem_cost_effectiveness.

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               # fig.height = 5,
               # fig.width = 8,
               fig.align = 'center')

```

```{r}
# Data preparation
source('master.R')

# Define theme for plotting
# Borrowed from https://rpubs.com/Koundy/71792
theme_publication <- function(base_size=14) {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.5, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
}

colors <- colorRampPalette(c('orange', 'red', 'darkgrey'))(3)

scale_fill_publication <- function(...){
  library(scales)
  discrete_scale("fill",
                 "Publication",
                 manual_pal(values = colors
                              # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)

}

scale_colour_publication <- function(...){
  library(scales)
  discrete_scale("colour",
                 "Publication",
                 manual_pal(values = colors
                            # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)

}

# Remove certain areas as per Ranjeeta's filtering
df <- 
  df %>%
  filter(disease == 'MALARIA',
         district != 'XAI-XAI CITY',
         district != 'MANHICA',
         district != 'MASSINGIR')

model_data <- df %>%
    filter(age_group == '0-4')

# Get a standardized time variable
model_data <-
  model_data %>%
  mutate(time = as.Date(paste0(year, '-01-01')) + (week*7)) %>%
  mutate(time = as.numeric(time)) %>%
  mutate(high_season = ifelse(month %in% 1:3, 1, 0))

# Generate a oct-dec lag
high_season <- model_data %>%
  filter(high_season == 1) %>%
  group_by(district, year) %>%
  summarise(cases = mean(pk)) %>%
  ungroup %>%
  group_by(district) %>%
  mutate(cases_lag = dplyr::lag(cases, n = 1, default = NA)) %>%
  ungroup %>%
  dplyr::select(-cases)
model_data <-
  left_join(model_data,
            high_season)
# Get 1 week lags too
model_data <-
  model_data %>%
  arrange(date) %>%
  group_by(district) %>%
  mutate(cases_lag1 = dplyr::lag(pk, n = 1, default = NA),
         cases_lag2 = dplyr::lag(pk, n = 2, default = NA))

# Filter away some stuff
model_data <- 
  model_data %>%
  filter(week < 52) %>%
  filter(date >= '2013-10-01')

# Get data with dummies
model_formula <- lm(pk ~  high_season + year + time, data = model_data)
model_matrix <- data.frame(model.matrix(model_formula)) %>%
  dplyr::select(-`X.Intercept.`) %>%
  mutate(pk = model_data$pk,
         district = model_data$district,
         itn_coverage = model_data$itn_coverage) %>%
  mutate(district_name = district) %>%
  mutate(district = as.numeric(factor(district)),
         cases_lag = model_data$cases_lag,
         cases_lag1 = model_data$cases_lag1,
         cases_lag2 = model_data$cases_lag2)

left <- 
  expand.grid(district = sort(unique(model_matrix$district)),
              time = sort(unique(model_matrix$time)))
right <- 
  model_matrix %>%
  group_by(district, time) %>%
  tally
balanced <- left_join(left, right)
model_matrix <-
  left_join(left, model_matrix)

```

# Map

```{r, fig.width = 8, fig.height = 5}
library(cism)
library(sp)
library(tidyverse)
library(broom)
library(maptools)
library(RColorBrewer)
library(ggrepel)

data <- moz3
data <- data[data@data$NAME_1 %in% c('Maputo', 'Gaza'),]
data@data$id <- paste0(data@data$NAME_1,
                       '|',
                       data@data$NAME_2)
data <- tidy(data, region = 'id')
data$province <- 
  unlist(lapply(strsplit(data$id, '|', fixed = TRUE), function(x){x[1]}))
data$district <- 
  unlist(lapply(strsplit(data$id, '|', fixed = TRUE), function(x){x[2]}))

label_df <- data %>%
  group_by(district) %>%
  summarise(long = mean(long),
            lat = mean(lat))

# Get centroids of provinces
centroids_p <-
  moz1@data %>%
  bind_cols(coordinates(moz1) %>% data.frame() %>% rename(x = X1, y = X2))

g1 <- ggplot() +
  theme_publication() +
  geom_polygon(data = cism::moz0_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = 'white',
               color = 'black',
               size = 0.2) +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = province),
               alpha = 0.5) +
  coord_map() +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'black',
               fill = NA,
               alpha = 0.6,
               size = 0.2) +
  scale_fill_manual(name = 'Province',
                    values = grey(c(0.05, 0.35))) +
  # geom_label_repel(data = label_df,
  #            aes(x = long,
  #                y = lat,
  #                label = district),
  #            label.padding = unit(0.05, "lines"),
  #            size = 3,
  #            alpha = 0.9) +
  geom_label_repel(data = centroids_p %>%
                     filter(NAME_1 %in% c('Gaza', 'Maputo')),
             aes(x = x,
                 y = y,
                 label = NAME_1),
             label.padding = unit(0.05, "lines"),
             box.padding = unit(3, 'lines'),
             arrow = arrow(length = unit(0.02, "npc")),
             size = 7,
             alpha = 0.9,
             segment.alpha = 0.6,
             alpha = 0.8,
             min.segment.length = unit(0, "lines")) +
  labs(x = 'Longitude',
       y = 'Latitude') +
  ggthemes::theme_map() +
  coord_map() +
  theme(legend.position = 'none',
        legend.text = element_text(size = 13)) +
  theme(plot.title = element_text(size=13),
        plot.subtitle = element_text(size = 16))

g2 <- ggplot() +
  coord_map() +
  theme_publication() +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'black',
               fill = NA,
               alpha = 0.6,
               size = 0.2) +
  scale_fill_manual(name = 'Province',
                    values = grey(c(0.05, 0.35))) +
  # geom_label_repel(data = label_df,
  #            aes(x = long,
  #                y = lat,
  #                label = district),
  #            label.padding = unit(0.05, "lines"),
  #            size = 3,
  #            alpha = 0.9) +
  labs(x = 'Longitude',
       y = 'Latitude') +
  ggthemes::theme_map() +
  coord_map() +
  theme(legend.position = 'right',
        legend.text = element_text(size = 13)) +
  theme(plot.title = element_text(size=13),
        plot.subtitle = element_text(size = 16)) +
    geom_polygon(data = data %>%
                   filter(district == 'Magude'),
               aes(x = long,
                   y = lat,
                   group = group),
               alpha = 0.5,
               fill = grey(0.1)) +
    geom_label_repel(data = data %>%
                       filter(district == 'Magude') %>%
                       group_by(district) %>%
                       summarise(long = mean(long),
                                 lat = mean(lat)),
             aes(x = long,
                 y = lat,
                 label = district),
             label.padding = unit(0.05, "lines"),
             box.padding = unit(6, 'lines'),
             arrow = arrow(length = unit(0.02, "npc")),
             size = 7,
             alpha = 0.9,
             segment.alpha = 0.6,
             alpha = 0.8,
             min.segment.length = unit(0, "lines"),
             nudge_x = 2)

library(grid)
library(gridExtra)
g <- grid.arrange(g1, g2, top = textGrob("Geography of intervention and control areas", gp=gpar(fontsize=15, fontface = 2)), ncol = 2)

# Define functionfor siaving files
save_figs <- function(g){
  if(!dir.exists('figs')){
    dir.create('figs')
  }
  file_name <- paste0(Sys.time())
  ggsave(filename = paste0('figs/', file_name, '.pdf'), g)
  ggsave(filename = paste0('figs/', file_name, '.png'), g)

}
save_figs(g)
```

# Trends in weekly incidence

```{r}
plot_data <-  df %>%
           mutate(location = ifelse(district == 'MAGUDE',
                                    'Magude',
                                    'Control districts')) %>%
           group_by(date, age_group, location) %>%
           summarise(cases = sum(cases),
                     population = sum(population)) %>%
           ungroup %>%
           mutate(pk = cases / population * 1000) %>%
  mutate(location = factor(location,
                           levels = c('Magude',
                                      'Control districts'))) %>%
  mutate(age_group = paste0('Age ', age_group))

g <- 
  ggplot(data =plot_data,
         aes(x = date,
             y = pk)) +
  geom_line(aes(linetype = location,
                color = location),
            alpha = 0.8) +
  facet_wrap(~age_group) +
  labs(x = 'Week',
       y = 'Cases per 1,000',
       title = 'Weekly incidence') +
  scale_linetype_manual(name = '',
                        values = c(1,2)) +
  theme_publication() +
  scale_color_manual(name = '',
                     values = c(grey(0.1), 'black')) +
  geom_vline(xintercept = as.Date('2015-10-10'),
             col = 'black',
             alpha = 0.8) +
  geom_label(data = data.frame(x = as.Date('2015-10-10'),
                              y = 20,
                              label = 'Intervention\nstart'),
             aes(x = x,
                 y = y,
                 label = label),
             size = 2.5)
g
save_figs(g)

```

# Cases as a percentage of total period average, ages 0-4 years

```{r}

# Define date truncation
date_truncate <- function (date_object, level = c("month", "quarter", "year")) 
  {
    if (is.null(level)) {
      stop("You must provide a level argument of either \"month\", \"quarter\" or \"year\".")
    }
    date_object <- as.Date(date_object)
    if (sum(!is.na(date_object)) == 0) {
      return(date_object)
    }
    if (level == "month") {
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y-%m"), "-01"))
      return(return_object)
    }
    if (level == "quarter") {
      q_month <- (((((as.numeric(format(date_object, "%m"))) - 
                       1)%/%3) + 1) * 3) - 2
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y"), ifelse(nchar(q_month[!is.na(return_object)]) == 
                                                                                    2, "-", "-0"), q_month, "-01"))
      return(return_object)
    }
    if (level == "year") {
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y"), "-01-01"))
      return(return_object)
    }
  }

# Get a date helper for standardizing dates
wh <- create_date_helper()
wh$month <- as.numeric(format(wh$date, '%m'))

# MANHICA OPD ########################################
load('misc_data/opd_cleaned.RData')
opd_manhica <- opd %>%
  filter(malaria) %>%
  filter(date >= min(bes$date)) %>%
  mutate(age_group = ifelse(age < (5 * 12), '0-4',
                            ifelse(age >= (5 * 12), '5+',
                                   NA))) %>%
  dplyr::select(age_group, date) %>%
  mutate(year = as.numeric(format(date, '%Y')),
         month = as.numeric(format(date, '%m')),
         week = as.numeric(format(date, '%U')) + 1) %>%
  group_by(year, week, age_group) %>%
  summarise(cases = n()) %>%
  ungroup %>%
  mutate(district = 'MANHICA',
         source = 'OPD')
opd_manhica_under5 <-
  opd_manhica %>%
  filter(age_group == '0-4') %>%
  group_by(year, week, district, source) %>%
  summarise(cases = sum(cases, na.rm = TRUE))

# MAGUDE OPD ###########################################
# Read in old OPD Magude data, sent by Bea
old_opd <- readstata13::read.dta13('misc_data/opd_magude_old/survmag_rrs_dis.dta')
old_opd <-
  old_opd %>%
  dplyr::select(yr,
                week,
                mal_hf) %>%
  rename(year = yr,
         cases = mal_hf)

opd_magude_all <- old_opd
opd_magude_all$district <- 'MAGUDE'
opd_magude_all$source <- 'OPD'
rm(old_opd)

# Combine all sources
comparison_magude <-
  opd_magude_all %>%
  bind_rows(bes %>%
              filter(district == 'MAGUDE',
                     disease == 'MALARIA') %>%
              group_by(year, week, district) %>%
              summarise(cases = sum(cases, na.rm = TRUE)) %>%
              ungroup %>%
              mutate(source = 'BES')) %>%
  left_join(wh) %>%
  group_by(year, month, source, district) %>%
  summarise(date = first(date),
            cases = sum(cases, na.rm = TRUE))

comparison_manhica <-
  opd_manhica_under5 %>%
  bind_rows(bes %>%
              filter(district == 'MANHICA',
                     age_group == '0-4',
                     disease == 'MALARIA') %>%
              group_by(year, week, district) %>%
              summarise(cases = sum(cases, na.rm = TRUE)) %>%
              ungroup %>%
              mutate(source = 'BES')) %>%
  left_join(wh) %>%
  group_by(year, month, source, district) %>%
  summarise(date = first(date),
            cases = sum(cases, na.rm = TRUE))

g1 <- ggplot(data = comparison_magude,
       aes(x = date,
           y = cases,
           color = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Magude: all cases',
       x = 'Date',
       y = 'Cases') +
  theme_cism() +
  scale_color_manual(name = 'Source',
                     values = c('darkgreen', 'darkorange'))

g2 <- ggplot(data = comparison_manhica,
       aes(x = date,
           y = cases,
           color = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Manhica: 0-4 only',
       x = 'Date',
       y = 'Cases') +
  theme_cism() +
  scale_color_manual(name = 'Source',
                     values = c('darkgreen', 'darkorange'))

# Comparison manhica
x <- 
  comparison_manhica %>%
  ungroup %>%
  group_by(district, source) %>%
  mutate(avg = mean(cases)) %>%
  ungroup %>%
  mutate(y = cases / avg * 100)
g <- 
  ggplot(data = x %>%
  filter(date <= '2016-08-30'),
       aes(x = date,
           y = y, 
           linetype = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Cases as a percentage of total period average',
       subtitle = 'Manhiça district, ages 0-4 only',
       x = 'Date',
       y = 'Percent') +
  theme_publication() +
  scale_linetype_manual(name = '',
                     values = c(1, 2)) 
g
save_figs(g)
```

```{r}
# Zip up all figures
zip('figs/figures.zip',
    files = paste0('figs/', dir('figs')))
```

```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```