### 2017-09-06 reports/aes/aes_effectiveness.Rmd
update

---
title: "Short-term impact of a malaria elimination campaign in Southern Mozambique"
subtitle: Evidence from a quasi-experimental approach applied to routine surveillance data
author: "AES, Barcelona, 2017"
header-includes:
  - \widowpenalties 1 150
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    theme: "Berlin"
    colortheme: "seahorse"
    fonttheme: "structuresmallcapsserif"
    toc: true
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r, results = 'hide'}
# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html") 
# Define a theme
theme_maltem <- ggthemes::theme_hc
```

```{r}
source('../../master.R', chdir = TRUE)

tablify <- function(x, n = 5){
  DT::datatable(x,
                selection = 'none',
                escape = FALSE,
                options = list(#sDom  = '<"top">lrt<"bottom">ip',
                               pageLength = n, 
                               dom = 'tip'))
}
kablify <- function(x, size = 16){
  kable(x) %>% 
  kable_styling(bootstrap_options = c("striped",
  "hover",
  "condensed"),
  font_size = size,
  position = 'float_left')
}

# ## 
# 
# <img src="img/tree.jpg" alt="img/tree.jpg" style="width: 600px; align="center"/>

# # Synthetic control reproduction
# 
# # Get packages
# library(Synth)
# 
# # Prepare data
# gen_dummy <- function(condition){ifelse(condition, 1, ifelse(!condition, 0, NA))}
# 
# model_data <- 
#   df %>%
#   filter(age_group == '0-4') %>%
#   filter(disease == 'MALARIA',
#          district != 'XAI-XAI CITY',
#          district != 'MANHICA',
#          district != 'MASSINGIR') 
# 
# # Get a standardized time variable
# model_data <-
#   model_data %>%
#   mutate(time = as.Date(paste0(year, '-01-01')) + (week*7)) %>%
#   mutate(time = as.numeric(time)) %>%
#   mutate(high_season = ifelse(month %in% 1:3, 1, 0))
# 
# # Generate a oct-dec lag
# high_season <- model_data %>%
#   filter(high_season == 1) %>%
#   group_by(district, year) %>%
#   summarise(cases = mean(pk)) %>%
#   ungroup %>%
#   group_by(district) %>%
#   mutate(cases_lag = dplyr::lag(cases, n = 1, default = NA)) %>%
#   ungroup %>%
#   dplyr::select(-cases)
# model_data <-
#   left_join(model_data,
#             high_season)
# # Get 1 week lags too
# model_data <-
#   model_data %>%
#   arrange(date) %>%
#   group_by(district) %>%
#   mutate(cases_lag1 = dplyr::lag(pk, n = 1, default = NA),
#          cases_lag2 = dplyr::lag(pk, n = 2, default = NA))
# 
# # Filter away some stuff
# model_data <- 
#   model_data %>%
#   filter(week < 52) %>%
#   filter(date >= '2013-10-01')
# 
# # Get data with dummies
# model_formula <- lm(pk ~  high_season + year + time, data = model_data)
# model_matrix <- data.frame(model.matrix(model_formula)) %>%
#   dplyr::select(-`X.Intercept.`) %>%
#   mutate(pk = model_data$pk,
#          district = model_data$district,
#          itn_coverage = model_data$itn_coverage) %>%
#   mutate(district_name = district) %>%
#   mutate(district = as.numeric(factor(district)),
#          cases_lag = model_data$cases_lag,
#          cases_lag1 = model_data$cases_lag1,
#          cases_lag2 = model_data$cases_lag2)
# 
# left <- 
#   expand.grid(district = sort(unique(model_matrix$district)),
#               time = sort(unique(model_matrix$time)))
# right <- 
#   model_matrix %>%
#   group_by(district, time) %>%
#   tally
# balanced <- left_join(left, right)
# model_matrix <-
#   left_join(left, model_matrix)
# 
# # Fill the NAs
# model_matrix <-
#   model_matrix %>%
#   arrange(time) %>%
#   group_by(district) %>%
#   mutate(high_season = zoo::na.locf(high_season), 
#          year = zoo::na.locf(year),
#          pk = zoo::na.locf(pk),
#          itn_coverage = zoo::na.locf(itn_coverage),
#          district_name = zoo::na.locf(district_name))
# 
# model_matrix <- data.frame(model_matrix)
#   
# prior_time <- sort(unique(model_matrix$time))
# prior_time <- prior_time[prior_time <= as.numeric(as.Date('2015-08-01'))]
# prepared_data <- 
#   Synth::dataprep(foo = model_matrix,
#                   predictors = c('itn_coverage',
#                                  'cases_lag',
#                                  'cases_lag1',
#                                  'cases_lag2'),
#                   predictors.op = 'mean',
#                   dependent = 'pk',
#                   time.variable = 'time',
#                   unit.variable = 'district',
#                   treatment.identifier = which(sort(unique(model_data$district)) == 'MAGUDE'),
#                   controls.identifier = which(sort(unique(model_data$district)) != 'MAGUDE'),
#                   unit.names.variable = 'district_name',
#                   time.predictors.prior = prior_time,
#                   time.optimize.ssr = prior_time,
#                   time.plot = sort(unique(model_matrix$time)))
# 
# out <- Synth::synth(data.prep.obj = prepared_data)
# # synth_tab <- synth.tab(synth.res = out,
# #           dataprep.res = prepared_data,
# #           round.digit = 3)
# # # Predictor weights
# # out$solution.v
# # # Unit weights
# # out$solution.w
```

# Introduction

## The burden of malaria

- 2015: 212M cases; 429K deaths (global). 
- Mozambique among ten highest burden countries in world. 
- 29% reduction since 2010.  
- Now, focus on _elimination_ and _eradication_.

## How to eliminate

- Mass drug administration (MDA).  
- Other, more time-tested strategies (ITNs, IRS).
- The need for operational research.

## It's hard to evaluate elimination strategies

- Difficult to measure impact of interventions in operational research.  
- OR is usually not experimental, not randomized, not controlled.
- Before vs. after and case vs. control approaches are useful, but biased.

## Why evaluate?

- The WHO explicitly supports “the need for more research on the optimum methods of implementing MDA programs, promoting community participation and compliance with treatment, and evaluating their effectiveness.

- As elimination initiatives imply both strong financial/economic efforts and a high risk of failure, their affordability, sustainability and cost-effectiveness are among the main concerns of funders, stakeholders and policymakers

- Evidence of the causal impact of malaria elimination initiatives is a pre-requisite for evaluating their cost-effectiveness. Finally, rigorous impact evaluations of malaria elimination programs are essential for sustained financial and political support.

## How to evaluate?

- Since we don't have an experiment, we need quasi-experimental approaches.  
- The synthetic control method (SCM).  
- We want to construct a _realistic_ counterfactual telling us _what would have happened_ without the campaign.

## What are we evaluating, exactly?

- Malaria elimination campaign in Magude.  
- Deployed during rainy seasons of 2015-6 and 2016-7. 
- Consisted of MDA, IRS, and surveillance. 
- ITN coverage was already high from 2014.

# Methods

## Data

- Mozambique’s Boletim Epidemiológico Semanal (BES) of the NMCP.  
- Weekly clinical cases, by age group, by health facility, for each of a few high risk diseases
, the epidemic disease reporting system used by the NMCP. 
- Population estimates from the National Statistical Institute (INE).

## Identification strategy

- Treatment group is district of Magude, Maputo province. 
- Control group consists of the other districts in Maputo and Gaza province. 
- The selection of control districts from locations around Magude is to ensure treatment and control groups have similar epidemiological characteristics.

## Co-variates

Our analysis uses variables as predictors of malaria incidence:

- LLINs

- Average weekly temperature

- Weekly rainfall

## Rainfall

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

g <- ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = precipitation,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = precipitation,
           group = district)) +
  labs(x = 'Date',
       y = 'Milimeters',
       title = 'Precipitation: Magude vs. other districts',
       subtitle = 'Monthly average precipitation, Magude in red') +
  theme_maltem()
g 
```

## Temperature

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = temp,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = temp,
           group = district)) +
  labs(x = 'Date',
       y = 'Degrees (celcius)',
       title = 'Average temperature: Magude vs. other districts',
       subtitle = 'Monthly average, Magude in red') +
  theme_maltem() +
  xlim(as.Date('2014-01-01'),
               max(x$date))
```

## Difference in differences

- We _could_ compare changes in malaria incidence to other districts using a difference-in-differences (DD) model. 
- Often used to estimate treatment effects in such settings by contrasting changes in outcome pre- and -post intervention, for the treatment and control groups.
- Eliminates confounding due to time-invariant differences between treatment and control groups, arising from unobservable factors that might influence malaria incidence. 
- Assumes that any macro shocks such as time effects are common to the groups under evaluation.

## Why we can't use difference in differences

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

g1 <- ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x2 <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  group_by(date) %>%
  summarise(magude_to_other = p[place == 'Magude'] / 
              p[place == 'Other districts'] * 100)

g2 <- ggplot(data = x2,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  geom_hline(yintercept = 100,
             col = 'darkred',
             alpha = 0.6,
             lty = 2) +
  labs(x = 'Date',
       y = 'Percent') +
  labs(title = 'Magude incidence as % of other districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))
Rmisc::multiplot(g1, g2, cols = 2)
```

## Why not DiD?

- Magude shows a downward trend in incidence over time, even prior to the elimination campaign.
- Lack of a visible parallel trend between treatment and control groups

## SCM

- We employ the synthetic control method (SCM).  
- Weighted combination of potential control districts, the synthetic control, using the districts that best approximate the most relevant characteristics of malaria incidence in Magude before the intervention. 
- Characteristics include pre-intervention malaria incidence during rainy season, temperature and and rainfall.
- Good match based on RMSE.

## Advantages of SCM

- Makes explicit the relative contribution of selected control districts to the counterfactual; the similarities between Magude and the synthetic control can be evaluated by comparing pre-intervention outcomes and predictors.  
- Allows the effects of unobserved factors on malaria incidence to vary with time, relaxing the parallel trend assumption of DD discussed earlier.

## Limitations of SCM

- Does not allow assessing the significance of the results using commonly used (large-sample) inferential techniques. 
- Instead, we use placebo experiments based on permutation techniques to make inference:  sequentially applying the SCM assuming each control district was “treated” to generate a distribution of placebo effects, and compare Magude's effect with these distributions.

# Results

## All districts' incidences (aggregated age groups)

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_line(alpha = 0.8,
            color = 'darkred') +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  facet_wrap(~district) +
  theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))
```

## All districts' incidences (separated by age groups)

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, age_group, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = age_group,
           color = age_group)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('darkorange', 'darkgreen')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~district) +
  # theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))
```

## Magude vs. other districts

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem()
```

## Magude vs. other districts (separated by age group)

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place, age_group) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~age_group)
```

## Our weighted synthetic control

```{r}
uw <-
  df %>%
  filter(age_group == '0-4') %>%
  filter(disease == 'MALARIA',
         district != 'XAI-XAI CITY',
         district != 'MANHICA',
         district != 'MASSINGIR') %>%
  group_by(district) %>%
  tally
uw$w.weight <-
  c(0.00,
    0.22,
    0.00,
    0.19,
    0.01,
    0.00,
    0.00,
    0.27,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.07,
    0.00,
    0.00,
    0.24
    )
uw$age_group <- '0 to 4'

older <- uw
older$age_group <- '5+'
older$w.weight <- 
  c(0.00,
    0.00,
    0.00,
    0.00,
    0.10,
    0.00,
    0.00,
    0.26,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.35,
    0.29
)
uw <- bind_rows(uw, older)
# Overwrite with Ranjeeta's weights!
ggplot(data = uw,
       aes(x = district,
           y = w.weight)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkorange') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'District',
       y = 'Weight',
       title = 'Weights by district') +
  ggthemes::theme_hc() +
  facet_wrap(~age_group,
             nrow = 2)
```

## Is it a good match?

```{r}
# Copy of Ranjeeta's table
x <- dplyr::data_frame(Indicator = c('Incidence, rainy season 2015',
                              'Incidence, rainy season 2014', 
                              'LLIN coverage',
                              'Pre-intervention precipitation',
                              'Pre-intervention temperature',
                              'Post-intervention precipitation'),
                `Magude 0-4` = c(3.65, 8.87, 24.82, 1.71, 24.52, 1.84),
                `Synthetic Magude 0-4` = c(5.1, 6.93, 24.77, 1.74, 24.51, 1.82),
                `Magude 5+` = c(4.02, 8.53, 24.83, 1.7, 24.52, 1.84),
                `Synthetic Magude 5+` = c(4.12, 6.29, 25.22, 1.71, 24.49, 1.83))
```

\begin{tiny}

```{r}

knitr::kable(x, format = 'latex')
```

\end{tiny}

## 0-4 year-olds: Magude vs synthetic control

<!-- <img src="img/synthetic1.png" alt="img/synthetic1.png" style="width: 800px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/synthetic1}

## 5+ year-olds: Magude vs synthetic control

<!-- <img src="img/synthetic2.png" alt="img/synthetic2.png" style="width: 800px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/synthetic2}

## Effect of interevention

Relative to the synthetic control, the average treatment effect during the first post-intervention malaria season was a reduction in weekly incidence by 4 and 2 cases per 1,000 in those aged 0-4 and 5 or older, respectively. During peak malaria season (January - March), 1,939 cases were prevented in 2016, and 3,405 in 2017.

<!-- <img src="img/ranjeeta_table.png" alt="img/ranjeeta_table.png" style="width: 600px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/ranjeeta_table}

This represents a 91.29% and 81.3% reduction in the incidence of malaria relative to the hypothetical control.

# Discussion

## Our findings

-  Across age groups during the peak malaria season (January-March) incidence declined by 91.29% and 81.3% in 2016 and 2017, respectively.  
- Has not achieved zero cases.  
- Unsure, at this point, to what extent we can attribute cases to importation or other causes.

## Comparison with other findings

- Similar findings obtained in a cluster randomized controlled trial (RCT) estimating the effectiveness of MDA with DHA/PPQ in Zambia.

- In that study, while the number of malaria cases significantly declined after MDA in low-transmission areas, drop was not significant in areas of high transmission (ie, Magude).

- Also consistent with the predictions of a consensus study that put together 4 models focusing on the impact of MDA with a drug having similar characteristics as DHA/PPQ.

- The latter predicted an immediate (but transient w/o other interventions) decline in malaria prevalence after MDA.

## Quasi-experimental methods in similar contexts

- Quasi-experimental approaches have been applied in other elimination campaigns.

- Barofsky et al. used SCM as an approach for the identification of the impact of a malaria campaign in a district in Uganda.   
- These studies are from 50s-60s and look at economic indicators, such as primary school completion or households consumption.   
- To our knowledge, this is the first quasi-experimental evaluation of an ongoing elimination initiative using routine national surveillance data on health indicators.

## Comparison with RCT

- Randomized control trials (RCTs) are the gold standard for impact evaluations.   
- However, the implementation of RCTs may imply financial, practical and ethical challenges.  
- Furthermore, RCTs have debatable usefulness in identifying intervention feasibility and implementation bottlenecks, which are particularly important in disease elimination contexts.
- Quasi-experimental approaches are particularly useful when applied to routinely collected (administrative/observational) data.

## Scaling up the method?

- Given the focus over the last decade on malaria elimination with both large and small scale campaigns across the globe – quasi-experimental approaches offer a rigorous and inexpensive alternative to randomized-control trials.

- Observational data is widely (and often publicly) available and true experimental approaches are often not feasible:  quasi-experimental approaches like the SCM are an excellent alternative for the evaluation of malaria elimination initiatives.

- Approach is not only suitable for the estimation of the impact of the intervention in question, but also generalizable to malaria elimination campaigns globally.

## Limitations

1. We rely on surveillance malaria cases reported at the health facility level, which is an incomplete and imperfect proxy for true malaria infection incidence since health-seeking behavior is especially low in low-income countries settings and may be different across geography and time.

2. The health facility registries from which we gathered data are themselves imperfect; we have no mechanisms by which to validate the entirety of the data. Still, BES data offers the best country-wide comparison of malaria trends at a granular level (weekly and by district).

3. Even with quasi-experimental data, we still can't fully control for abnormalities... (see next slide)

## 2017 has been anything but normal

```{r}
x <- 
  df %>%
  group_by(year, month) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(year = factor(year))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = month,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  geom_point() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  labs(x = 'Month',
       y = 'Average incidence',
       title = 'Incidence by month, all districts',
       subtitle = 'Wide variability in monthly incidence') +
  theme_maltem()

```

## Implications

- The strategy reduced malaria significantly.   
- Effectiveness quantified accurately enough to allow for granular cost-effectiveness.   
- Synthetic control method useful and generalizable for the estimation of similar initiatives’ effectiveness.   
- Policy-makers, researchers, and those actively engaged in or planning disease elimination initiatives, can benefit from both the methods and findings from this study.

## Acknowledgments

- NMCP for releasing BES data.  
- CISM for support.  
- Bill and Melinda Gates Foundation and “La Caixa” Foundation for funding MALTEM.   
- JB is funded by the Trans Global Health doctorate program.